### 2019-01-14 submission/manuscript.Rmd
Fix legend placement

---
title: ''
csl: mbio.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  word_document: default
geometry: margin=1.0in
bibliography: references.bib
---

@@

The authors would like to thank all the study participants in ERIN whose samples were utilized. We also would like to thank Judy Opp and April Cockburn for their effort in sequencing the samples as part of the Microbiome Core Facility at the University of Michigan. Additional thanks to members of the Schloss lab and Dr. Marcy Balunas for reading earlier drafts of the manuscript and providing helpful critiques. Salary support for Marc A. Sze came from the Canadian Institute of Health Research and NIH grant UL1TR002240. Salary support for Patrick D. Schloss came from NIH grants P30DK034933 and 1R01CA215574.

@@

**Figure 1: XXXXXXXX.**

**Figure S1: XXXXXXXX.**

>>>>

---
title: ''
csl: mbio.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
bibliography: references.bib
---

@@

The authors thank the study participants in ERIN whose samples were utilized. We also would like to thank Judy Opp and April Cockburn for their effort in sequencing the samples as part of the Microbiome Core Facility at the University of Michigan. Additional thanks to members of the Schloss lab and Dr. Marcy Balunas for reading earlier drafts of the manuscript and providing helpful critiques. Salary support for Marc A. Sze came from the Canadian Institute of Health Research and NIH grant UL1TR002240. Salary support for Patrick D. Schloss came from NIH grants P30DK034933 and 1R01CA215574.

@@

## Figure Lengends

**Figure 1. The error rate of assembled sequence reads increases with the number of rounds of PCR used and follows the relative error rates provided by the manufacturers; however, much of this error is mediated by denoising.**

**Figure 2. The fraction of all denoised sequences that were identified as being chimeric increases with the number of rounds of PCR used and varied between polymerases.** (A) Sequencing of a mock community allowed us to identify the total fraction of sequences that were chimeric as well as the sensitivity and specificity of UCHIME to detect those chimeras. (B) Sequencing of four human stool samples after using one of five different polymerases again demonstrated increased rate of chimera formation with increasing number of rounds of PCR and variation across polymerases.

**Figure 3. The relative abundances of reads mapped to reference sequences differed subtly from the expected relative abundances as determined by shotgun metagenomic sequencing.** Bias did not increase with number of rounds of PCR or vary by polymerase or the guanine and cytosine content of the fragment. The expected relative abundance of each organism is indicated by the horizontal gray line. The percentage of bases that were guanines or cytosines within the V4 region of the 16S rRNA genes in each organism is indicated by the number in the lower left corner of each panel.

**Figure 4. Despite evidence of subtle PCR bias at the genome level, there was significant evidence of bias using community-wide metrics that grew with the number of rounds of PCR when using a mock community.** (A) With the exception of data collected using the Kappa polymerase, the OTU richness and Shannon diversity values increased with number of rounds of PCR and the inclusion of residual sequencing errors and chimeras. The horizontal black line indicates the expected richness and diversity if there were no bias, errors, or chimeras. (B) Relative to the mock community sampled after 25 rounds of PCR, the Bray-Curtis distance to the communities sampled after 30 and 35 rounds of PCR increased for all polymerases. (C) The variation between samples collected after 20, 25, 30, and 35 rounds of PCR for the five polymerases demonstrated a significant change in the community driven by the number of rounds of PCR and the polymerase used.

**Figure 5. Sequencing of human stool samples indicated clear increase in bias with number of rounds of PCR, however, the bias appeared to be consistent within each sample.** (A) With the exception of data collected using the Kappa polymerase, the OTU richness and Shannon diversity values increased with number of rounds of PCR. (B) Relative to the stool communities sampled after 25 rounds of PCR, the Bray-Curtis distance to the stool communities sampled after 30 and 35 rounds of PCR was inconsistent and there was little difference in variation for data collected using the Kappa polymerase. (C) The variation between stool samples was larger than the amount of variation introduced by varying the number of rounds of PCR or polymerase.

**Figure 6. There is little evidence for PCR drift adversely impacting the results of amplicon studies.** The average distance between replicates of sequencing the same mock community or between the the human stool samples did not vary by number of rounds of PCR or by polymerase.

**Figure S1: With the exception of the seuqence data generated using the Kappa polymerase, the ratio of the two *Salmonella enterica* V4 sequences to a value was lower than the expected ratio of 6:1.** The two *S. enterica* V4 sequences differed by a single base.
